--- 
title: "Hands-on Machine Learning with R & Python"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: misk-data-science/misk-homl
description: "Hands-on Machine Learning with R & Python."
---

# Welcome {-}

Welcome to *Hands-On Machine Learning*.  This book provides hands-on modules for many of the most common machine learning methods to include:

- A proper modeling process
- Feature engineering
- Linear and logistic regression
- Regularized models
- K-nearest neighbors
- Random forests
- Gradient boosting machines
- Stacking / super learners
- And more!

You will learn how to build and tune these various models with R and Python packages that have been tested and approved due to their ability to scale well. However, our motivation in almost every case is to describe the techniques in a way that helps develop intuition for its strengths and weaknesses.  For the most part, we minimize mathematical complexity when possible but also provide resources to get deeper into the details if desired.


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Who should read this {-}

```{block, type='todo'}
TBD
```

This module makes a few assumptions of your established knowledge regarding your programming skills. Below are some important assumptions made and resources to read through if you feel inadequately prepared:

```{block, type='todo'}
Complete after rest of material is finalized.
```

| Assumptions                       | `r fontawesome::fa("python")` Resource | `r fontawesome::fa("r-project")` Resource   
| --------------------------------- | ------------- | ------------- |
| You should be familiar with...    |               |               | 
| You should be familiar with...    |               |               |
| You should be familiar with...    |               |               |
| You should be familiar with...    |               |               |

## What you need for this book {-}

```{block, type='todo'}
TBD
```

## Conventions used in this book {-}

The following typographical conventions are used in this book:

* ___strong italic___: indicates new terms,
* __bold__: indicates package & file names,
* `inline code`: monospaced highlighted text indicates functions or other commands that could be typed literally by the user,
* code chunk: indicates commands or other text that could be typed literally by the user

```{r, first-code-chunk, collapse=TRUE}
1 + 2
```

We will also refer to both R and Python throughout. When discussing one or the other we will typically label with the appropriate icons:

- Python: `r fontawesome::fa("python")`
- R: `r fontawesome::fa("r-project")`

In addition to the general text used throughout, you will notice the following code chunks:

```{block, type = "tip"}
Signifies a tip or suggestion
```

```{block, type = "note"}
Signifies a general note
```

```{block, type = "warning"}
Signifies a warning or caution
```

## Feedback {-}

Reader comments are greatly appreciated. To report errors or bugs please post an issue at https://github.com/misk-data-science/misk-homl/issues.

## Acknowledgments {-} 

```{block, type='todo'}
TBD
```
