---
title: "Multivariate Adaptive Regression Splines"
output: html_notebook
---

# Prerequisites

Packages:

```{r}
# Helper packages
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting

# Modeling packages
library(earth)     # for fitting MARS models
library(caret)     # for automating the tuning process
library(rsample)  # for sampling

# Model interpretability packages
library(vip)       # for variable importance
library(pdp)       # for variable relationships
```

Data:

```{r prereqs-data}
# ames data
ames <- AmesHousing::make_ames()

# split data
set.seed(123)
split <- initial_split(ames, strata = "Sale_Price")
ames_train <- training(split)
```


# Tuning parameters

MARS models have two tuning parameters:

1. nprune: the maximum number of terms in the pruned model (including the intercept)
2. degree: the maximum degree of interaction

```{r earth-tuning-params}
caret::getModelInfo("earth")$earth$parameters
```


# Implementation

```{r cv-mars}
# tuning grid
hyper_grid <- expand.grid(
  nprune = seq(2, 50, length.out = 10) %>% floor(),
  degree = 1:3
)

# perform resampling
set.seed(123)
cv_mars <- train(
  Sale_Price ~ ., 
  data = ames_train, 
  trControl = trainControl(method = "cv", number = 10),
  method = "earth", #<<
  tuneGrid = hyper_grid,
  metric = "RMSE"
  )

# best model
cv_mars$results %>%
  filter(
    nprune == cv_mars$bestTune$nprune,
    degree == cv_mars$bestTune$degree
    )
```

```{r cv-mars-plot, fig.height=5}
# plot results
plot(cv_mars)
```

# Feature importance

* Backwards elimination feature selection routine that looks at reductions in the GCV estimate of error as each predictor is added to the model. 
* This total reduction is used as the variable importance measure (`value = "gcv"`). 
* Can also monitor the change in the residual sums of squares (RSS) as terms are added (`value = "rss"`)

**Automated feature selection**

```{r mars-vip, fig.height=10, fig.width=15}
p1 <- vip(cv_mars, num_features = 40, geom = "point", value = "gcv") + ggtitle("GCV")
p2 <- vip(cv_mars, num_features = 40, geom = "point", value = "rss") + ggtitle("RSS")
gridExtra::grid.arrange(p1, p2, ncol = 2)
```


# Partial dependence plots

```{r pdp, fig.width=15, fig.height=3, warning=FALSE, message=FALSE}
# Construct partial dependence plots
p1 <- partial(cv_mars, pred.var = "Gr_Liv_Area", grid.resolution = 10) %>% 
  ggplot(aes(Gr_Liv_Area, yhat)) +
  geom_line()
p2 <- partial(cv_mars, pred.var = "Year_Built", grid.resolution = 10) %>% 
  ggplot(aes(Year_Built, yhat)) +
  geom_line()
p3 <- partial(cv_mars, pred.var = c("Gr_Liv_Area", "Year_Built"), 
              grid.resolution = 10) %>% 
  plotPartial(levelplot = FALSE, zlab = "yhat", drape = TRUE, colorkey = TRUE, 
              screen = list(z = -20, x = -60))
# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

